Capstone Movielens Report

### 2025-05-23 reports/templates/02.X.UM+model.template.Rmd
UMGYE Model, Support functions described.

\begin{equation}
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + \gamma(q) + \varepsilon_{i,j}
(\#eq:UMGYE-model)
\end{equation}

@@

::: {.noteblock data-latex=""}
The complete source code of the functions described in this section is available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) section of the [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) script on _GitHub_.
:::

#### [train_UM?_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM-effect.functions.R#L2) Function {#func.train_UM?_effect}
\

We use this function to build and train our model using the `train_set` dataset:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The function described above accepts the `lambda` parameter, which we will need later for the _Model Regularization_ method. We also use the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L63) function call, which we will need later for the *Regularization* techniques (for details, see the  [mean_reg](#func.mean_reg) function description in the Section [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report).
We have already explained that in the [UM? Effect Model Regularization] section. 
For now, we omit the `lambda` parameter, accepting its default value `lambda = 0`. Let's recall that in this case, the [mean_reg](#func.mean_reg) function is equivalent to the standard R function [base::mean](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean).
:::

#### `train_UM?E_effect.cv` Function {#func.train_UM?E_effect.cv}
\

We use the [train_UM?E_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/#L59) function to build and train our model using the `5-Fold Cross Validation` method. Below, we provide a slightly simplified version of the source code of that function:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
Here we use the function call [union_cv_results](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L432) to aggregate the _5-Fold Cross Validation_ method results (for details, see the [union_cv_results](#func.union_cv_results) function description in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report).
:::

#### `calc_UM?E_effect_MSE` Function
\

The source code of the function [calc_UM?E_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/#L) defined in the [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) script to calculate the _Mean Squared Error (MSE)_ of the _UM?E Model_ for the given _Test Set_ is provided below:
```{r eval=FALSE}
```

@@

#### `calc_UM?E_effect_MSE.cv` Function
\

The source code of the function [calc_UM?E_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/#L) defined in the [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) script to calculate the _5-Fold Cross Validation MSE_ result of the _UM?E Model_ is provided below:
```{r eval=FALSE}
```

#### `calc_UM?E_effect_RMSE` Function
\

The source code of the function [calc_UM?E_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/#L118) defined in the [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) script to calculate the _Root Mean Squared Error (RMSE)_ of the _UM?E Model_ for the given _Test Set_ is provided below:

#### `calc_UM?E_effect_RMSE.cv` Function { #func.calc_UM?E_effect_RMSE.cv }
\

The source code of the function [calc_UM?E_effect_RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/#L) defined in the [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/) script to calculate the _5-Fold Cross Validation **RMSE**_ result of the _UM?E Model_ is provided below:

@@

>>>>

\begin{equation}
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + \gamma(q) + \varepsilon_{i,j}
(\#eq:UM?E-model)
\end{equation}

@@

::: {.noteblock data-latex=""}
Some of the functions described in this section accepts the `lambda` parameter, which we will need later for the _Model Regularization_ method. We use the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L63) function call as well, which we will also need for the *Regularization* techniques to apply to our models (for details, see the  [mean_reg](#func.mean_reg) function description in the Section [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report).
We have already explained that in the [UME Model Regularization] section. 
For now, we omit the `lambda` parameter, accepting its default value `lambda = 0`. Let's recall that in this case, the [mean_reg](#func.mean_reg) function is equivalent to the standard R function [base::mean](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean).
:::

#### [calc_UM?_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L70) Function {#func.calc_UM?_effect}
\

This is a helper function that calculates the _UM? effect_ of the user ratings and is intended to be used in other functions and scripts (for instance in [train_UM?_effect](#func.train_UM?_effect)) described below.

##### Parameters
\

The function accepts the only parameter `date_general_effect`, which is a data structure returned by any of the following functions described above:

  - [calc_date_general_effect](#func.calc_date_general_effect);
  - [calc_date_general_effect.cv](#func.calc_date_general_effect.cv).

##### Source Code
\

The following is the source code of the function:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the [calc_UM?_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L70) function is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L1) section of the [UM?-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L70) script on _GitHub_.
:::

##### Return

A data structure containg information about _User+Movie+Genre+Year (UM?) Effect_.

#### [train_UM?_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L75) Function {#func.train_UM?_effect}
\

This function is used to train the current model using the _Train Set_ dataset passed in the `train_set` parameter.
The following is the source code of the function:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the [train_UM?_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L75) function is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L1) section of the [UM?-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L75) script on _GitHub_.
:::

##### Return

A data structure containg information about _User+Movie+Genre+Year (UM?) Effect_.

#### [train_UM?_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L85) Function {#func.train_UM?_effect.cv}
\

This function is used to train the current model using the _5-Fold Cross Validation_ method.
The following is the source code of the function:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the [train_UM?_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L85) function is also available in the [UM?E Model Support Functions](#L1) section of the [UM?-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L85) script on _GitHub_.
:::

##### Return

A data structure containg information about _User+Movie+Genre+Year (UM?) Effect_.

#### [calc_UM?_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L93) Function
\

The function calculates the _Mean Squared Error (MSE)_ of the _UM?E Model_ for the given _Test Set_ passed in the `test_set` parameter.

##### Parameters
\

  - **test_set:** A _Test Set_ dataset used for _UM?E Model_ validation;
  - **UM?_effect:** A data structure returned by any of the following functions described above:
    - [calc_UM?_effect](#func.train_UM?_effect);
    - [train_UM?_effect](#train_UM?_effect) functions.
    - [train_UM?_effect.cv](#train_UM?_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the function [calc_UM?_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L93) is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L93) script. 
:::

@@

#### [calc_UM?_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) Function
\

The function calculates the _5-Fold Cross Validation MSE_ result for the _UM? Effect_ data passed in the `UM?_effect` parameter.

##### Parameters
\

The function accepts the only parameter: `UM?_effect`, which is a data structure returned by the one of the following functions described above:

  - [calc_UM?_effect](#func.train_UM?_effect);
  - [train_UM?_effect](#train_UM?_effect) functions.
  - [train_UM?_effect.cv](#train_UM?_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}
calc_UM?_effect_MSE.cv <- function(UM?_effect){
  start <- put_start_date()
  UM?_effect_MSEs <- sapply(edx_CV, function(cv_fold_dat){
    cv_fold_dat$validation_set |> calc_UM?_effect_MSE(UM?_effect)
  })
  put_end_date(start)
  put_log1("Function: calc_UM?_effect_MSE.cv
Date (Year) Effect MSE values have been computed for the %1-Fold Cross Validation samples.", 
           CVFolds_N)

  mean(UM?_effect_MSEs)
}

```

::: {.noteblock data-latex=""}
The source code of the function [calc_UM?_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) script. 
:::

#### [calc_UM?_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L115) Function
\

The function calculates the _Mean Squared Error (MSE)_ of the _UM?E Model_ for the given _Test Set_ passed in the `test_set` parameter.

##### Parameters
\

  - **test_set:** A _Test Set_ dataset used for _UM?E Model_ validation;
  - **UM?_effect:** A data structure returned by any of the following functions described above:
    - [calc_UM?_effect](#func.train_UM?_effect);
    - [train_UM?_effect](#train_UM?_effect) functions.
    - [train_UM?_effect.cv](#train_UM?_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the function [calc_UM?_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L115) is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L115) script. 
:::

#### [calc_UM?_effect_RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) Function
\

The function calculates the _5-Fold Cross Validation MSE_ result for the _UM? Effect_ data passed in the `UM?_effect` parameter.

##### Parameters
\

The function accepts the only parameter: `UM?_effect`, which is a data structure returned by the one of the following functions described above:

  - [calc_UM?_effect](#func.train_UM?_effect);
  - [train_UM?_effect](#train_UM?_effect) functions.
  - [train_UM?_effect.cv](#train_UM?_effect.cv).

##### Source Code
\

The source code of the function is provided below:
```{r eval=FALSE}

```

::: {.noteblock data-latex=""}
The source code of the function [calc_UM?_effect_RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) is also available in the [UM?E Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UM?-effect.functions.R#L119) script. 
:::

@@

## Additional Function Templates
\

#### [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L) Function {#func.}
\

::: {.noteblock data-latex=""}
Note
:::

#### [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L) Function {#func.}
\

::: {.noteblock data-latex=""}
Note
:::

#### [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L) Function {#func.}
\

::: {.noteblock data-latex=""}
Note
:::

#### [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L) Function {#func.}
\

::: {.noteblock data-latex=""}
Note
:::

#### [](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L) Function {#func.}
\

::: {.noteblock data-latex=""}
Note
:::
